---
title: "Day 3: Synthesis and Creation"
author: "Cleo Falvey & Filip Stefanovic"
date: "8/2/2020"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
```

# Welcome back!

Today we're going to be synthesizing and implementing Tidyverse! If today's RMarkdown looks a little bit empty to you, that's because we're going to be all making it together!

However, I'll reupdate the GitHub at the end so everyone has working code by the end of the day. In addition, to see how I prepared for this workshop (mainly because live coding demos stress me out), check out the "bonus_content.Rmd" file, Day 3.

# Putting it all together

Today is going to be a little different than usual. We're going to all code together! I'm going to start everyone off with a little bit of readr/dplyr, but then we're all going to split out into breakout groups and make graphs. It may be helpful to have yesterday's knitted .html file on hand, so you can remember the dplyr and ggplot2 syntax.

### Reading Data

Read in your data here. The data is a file in the working directory called "Iris.csv". It's about different types of flowers.

```{r}
# read in your data here.
```

### ggplot2

```{r}
# what plots do we think are interesting? this is an open-ended question; there is no right answer! however, always think about the data story we want to tell.

```

### dplyr

```{r}
# we've done some exploratory plotting. is there any way we need to change the shape or mutate our data files? what could be interesting?

```

### ggplot2 again (after dplyr)

```{r}
# if necessary, have we plotted our brand new dataset?

```

### Modelling

```{r}
# let's create a linear model of the relationship between two variables we think are interesting.
```

# Miscellaneous Takeaways: Day 3

- One of the most common bugs is actually using %>% for ggplot2 and + for dplyr. We ALWAYS use %>% for dplyr and + for ggplot2!
- See the "bonus_content.Rmd" file for more analyses, tips, and tricks that are already ready-to-run (i.e., hit "knit" to run it yourself, or open up the .html file!)

### Enjoy your R journey and have a great rest of your summer! Thanks much for participating! If you ever have an R question feel free to reach me at cleo.falvey001@umb.edu :)