---
title: "Iris Analyses"
author: "Cleo Falvey & Filip Stefanovic"
date: "8/6/2020"
output: html_document
---

# Libraries Loading

```{r}
library(tidyverse)
```


# Reading in Data

```{r}
iris <- read_csv("Iris.csv")
```

# Petal Length, Petal Width

```{r}
ggplot(iris, aes(x=Petal.Length, y=Petal.Width, color=Species)) +
  geom_point() +
  labs(x="Petal Length", y="Petal Width", title="Petal Length vs. Petal Width, Sorted by Species")
```


# Sepal Length, Boxplot

```{r}
ggplot(iris, aes(x=Sepal.Length, group=Species, y=Species, fill=Species)) +
  geom_boxplot()
```

# Virginica is consistently the largest flower. One of the things that we might consider is examining the ratio of flower sizes. Sometimes that changes things!

### To do this we are going to use dplyr & ggplot2.

```{r}
iris %>%
  mutate(petal_ratio = Petal.Length/Petal.Width) %>%
  ggplot(aes(x=petal_ratio, y=Species, group=Species, fill=Species)) +
  geom_boxplot()
```


# These are just a few ideas I wanted to come prepared with to the R workshop. Feel free to come up with your own!